---
title: "Project_Machine_Learning"
author: "Sandra Rairán"
date: "9/7/2020"
output: html_document
---

# Create a report describing how you built your model
#### RF library("caret") very slow then use library("randomForest")
1. read data  (pml-training.csv) and (pml-testing.csv)
2. check classes with unique(training$classe) [1] "A" "B" "C" "D" "E"
3. use complete.cases() to print a logical vector that indicates complete and missing rows (i.e. rows without NA).
and The data were initially inspected. A number of variables appeared to have missing variables e.g., max_roll_belt.
4. The na.strings argument is for substitution within the body of the file, that is, matching strings that should be replaced with NA.
5. Delete columns with all missing values.
6.delete the first 7 columns are not necessary for prediction with rf.
7. the dimensions of the new df
8. Data Slicing
Cross Validation
60% traning
40% testing
9. Random Forecast for prediction
10. Importance of variables
11. This graph shows the error vs. the number of trees. You can see how the error decreases as the number of trees increases.
12. confusionMatrix
13. predict final:
 1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 19 20 
 B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 
Levels: A B C D E

## Expected out-of-sample error

Out-of-bag (OOB) error, also called out-of-bag estimate, is a method of measuring the prediction error of random forests,

# Data Raw Loading and cleaning
```{r libraries,  results='hide'}
library("randomForest")
library("dplyr")
library("caret")
```


```{r data, echo=TRUE, warning = FALSE}

training = read.csv("pml-training.csv", 
header = TRUE)
testing<-read.csv("pml-testing.csv", header=TRUE)

# The data were initially inspected. A number of variables appeared to have missing variables e.g., max_roll_belt.
str(training)
## training has 160 variables of 19622 rows. 
dim(testing)
dim(training)

unique(training$classe)
#We can use complete.cases() to print a logical vector that indicates complete and missing rows (i.e. rows without NA). 
dfmissing <- sum(complete.cases(training))
dfmissing ## 406

## The na.strings argument is for substitution within the body of the file, that is, matching strings that should be replaced with NA
trainingdf <- read.csv("pml-training.csv", na.strings=c("NA", ""))
testingdf <- read.csv('pml-testing.csv', na.strings=c("NA", ""))
dim(trainingdf)
dim(testingdf)

dftrainingdf <- sum(complete.cases(trainingdf))
dftrainingdf

# Delete columns with all missing values
trainingdf<-trainingdf[,colSums(is.na(trainingdf)) == 0]
testingdf <-testingdf[,colSums(is.na(testingdf)) == 0]


## delete the first 7 columns are not necessary for prediction with rf
trainingdf   <-trainingdf[,-c(1:7)]
testingdf <-testingdf[,-c(1:7)]

## the dimensions of the new df
dim(trainingdf) ## 19622    53
dim(testingdf) ## 20 53
```

# Data Slicing
Cross Validation
60% traning
40% testing
```{r Slicing, echo=TRUE, warning = FALSE}
subsamples <- createDataPartition(y=trainingdf$classe, p=0.6, list=FALSE)
subTraining <- trainingdf[subsamples, ] 
subTesting <- trainingdf[-subsamples, ]

dim(subTraining)
dim(subTesting)
```


# Random Forecast for prediction

```{r randomForest1, echo=TRUE, warning = FALSE}
# Fit model
## mtry: Number of variables randomly sampled as candidates at each split.
subTraining$classe = factor(subTraining$classe )
model_rf <- randomForest(classe ~. , data=subTraining, method="class",mtry = 15, importance = TRUE)
prediction_rf <- predict(model_rf, subTesting, type = "class")

# Test results on subTesting data set:
model_rf
```
Out-of-bag (OOB) error, also called out-of-bag estimate, is a method of measuring the prediction error of random forests,

Importance of variables
```{r Importance, echo=TRUE}

importance(model_rf)
```
This graph shows the error vs. the number of trees. You can see how the error decreases as the number of trees increases.
```{r pltot, echo=TRUE}
plot(model_rf)

```



```{r randomforestcm, echo=TRUE,warning = FALSE}

confusionMatrix(table(prediction_rf, subTesting$classe))

```



```{r predictfinal1, echo=TRUE,warning=FALSE}

predictfinal <- predict(model_rf, testingdf, type="class")
predictfinal
```


